---
title: "stl-lead-inequity"
format: html
editor: visual
---

## Task 1

```{r}
library(tidyverse)
library(janitor)
library(here)
```

```{r}
stl_lead <-read_csv(here::here("data","stl_lead.csv")) %>% 
  janitor::clean_names()
```

## Summary Statistics

```{r}
summary(stl_lead)
head(stl_lead)
```

## Visualizations

```{r}
ggplot(data = stl_lead, aes(x = total_pop, y = total_pop_moe))+ geom_point(aes(color = poverty_tot_moe))
```

-   In a new code chunk, from `stl_lead` create a new data frame called `stl_lead_prop` that has one additional column called `prop_white` that returns the percent of each census tract identifying as white (variable `white` in the dataset divided by variable `totalPop`, times 100). You may need to do some Googling. Hint: `dplyr::mutate(new_col = col_a / col_b)` will create a new column `new_col` that contains the value of `col_a / col_b`

```{r}

stl_lead_prop <- dplyr::mutate(stl_lead,prop_white = (stl_lead$white/stl_lead$total_pop)*100)
```

## Create a scatterplot

-   In a new code chunk, create a scatterplot graph of the percentage of children in each census tract with elevated blood lead levels (`pctElevated`) versus the percent of each census tract identifying as white.

-   Customize by updating several aesthetics (e.g. size, opacity (see `alpha =`), color, etc.)

-   Store the scatterplot as `stl_lead_plot`

-   Have the scatterplot returned in the knitted html - customize the size that it appears when knitted

-   Also save a .png of the scatterplot to `figs`, with dimensions of (6\" x 5\") (width x height)

-   In text above or below the scatterplot, write 1 - 2 sentences describing the overall trend that you observe from your graph

```{r}
stl_lead_plot<-ggplot(data = stl_lead_prop, aes(x = prop_white, y = pct_elevated))+
  geom_point(alpha = 0.5, shape = 16, aes(color = white_moe))

stl_lead_plot

ggsave(here::here("figs", "stl_lead_plot.png"), height = 5, width = 6)
```

The scatterplot shows a slight negative linear trend of of pct elevations and proportion of white population.

## Histogram Plot

-   Create a histogram of only the `pctElevated` column in the data frame (remember, this will only take one variable - the frequency is calculated for you by `geom_histogram`)

-   Customize the fill, color, and size aesthetics - test some stuff! Feel free to make it awful.

-   Once you\'ve played around with customization, export the histogram as a .jpg to the `figs` folder

-   Make sure the histogram also shows up in your rendered html

```{r}
stl_lead_hist <- ggplot(data = stl_lead_prop, aes(x = pct_elevated))+
  geom_histogram(color = "blue", fill = "cyan")

stl_lead_hist

ggsave(here::here("figs", "stl_lead_hist.png"), height = 5, width = 6)
```
